---
title: "Quickstart"
output:
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning=FALSE, 
  message=FALSE
)

print.nicedf<-function(x){
  knitr::kable(x,format = "html")
}

```

```{r setup}
library(clog)
library(dplyr)
```

## Overview





## Input Example

```{r,include=FALSE}
my_cleaninglog<-data.frame(rowids = c("x8749f4", "f839jzz", "hla84s"),
                           dataset_variables = c("var_A","var_A","var_B"),
                           changed_to = c("maybe","no", NA),
                           found_via = c("manual inspection","manual inspection","manual inspection"),
                           stringsAsFactors = F)

my_dataset<-data.frame(uuid= c("x8749f4", "f839jzz", "hla84s","d88f30","bkoe4u"),
                   var_A = c("yes","yes","no","no","yes"),
                   var_B = c(10,20,30,40,50),
                   enumerator_id = c("Z","Z","X","Y","Y"),
                   stringsAsFactors = F)

```

### cleaning log

A cleaning log is a table in which each row represents a change to one specific value in a dataset. It must contain:

- a column with the names of the variables in the dataset that each change relates to
- a name for the type of change (for example the name of a check that originally flagged the value to be changed)
- a column with an ID matching a unique value in a variable in the dataset

Example:

```{r}
my_cleaninglog
```

### Dataset

For this to be usable, we need to know in addition **the name of the variable in the dataset (in quotes!) that matches the ids noted in the cleaninglog**

Example dataset: 
```{r}
my_dataset

```

Here, the "uuid" column in the dataset is unique and can be matched with the values in the "rowids" of the cleaning log.

## Applying a Cleaning log

### creating a standardised cleaninglog object

First we need to tell R that we are dealing with a cleaninglog, and what relates to what in the components mentioned above. We create a 'standardised' cleaninglog object with `cleaninglog()`:

```{r}
my_cleaninglog <- cleaninglog(ids = my_cleaninglog$rowids,
                              variables = my_cleaninglog$dataset_variables,
                              new_values = my_cleaninglog$changed_to,
                              name = my_cleaninglog$found_via,
                              data_id_column_name = "uuid")
```

Note that each parameter takes a vector of the same length; here we take them from a cleaninglog dataset that could for example come from a csv file. Only the parameters `data_id_column_name` must be the _name_ of a variable in the dataset - not the cleaninglog itselft - as a string. For details, run `?cleaninglog`.

Here's what we get:

```{r}
my_cleaninglog
```

The cleaninglog didn't change much from our original cleaninglog dataframe, but has standardised column headers, and secretly remembers which column in the dataset it should look for the row ids.

### Modify data based on cleaning log

Now we can change the dataset based on the cleaninglog:

```{r}

cleaned_data <- clog_clean(my_dataset,my_cleaninglog)
```
see the difference:
```{r}
my_dataset

cleaned_data
```

## Cleaned data memory

The cleaned dataset "remembers" the changes made, so the cleaned data can be reversed and can reproduce the cleaning log(s) that was used to modify it:

```{r}
clog_reverse(cleaned_data)
```


```{r}
clog_get_cleaninglog(cleaned_data)
```

## Moving information between cleaninglog and data

Apart from modifying the dataset, sometimes we want to move information between the cleaning log and the data (usually as part of the follow up process).

### Adding data to a cleaninglog

You can append variables from the data to the cleaninglog - often useful to follow up on values based on some metadata:

```{r}
my_dataset %>% clog_data_to_cleaninglog(my_cleaninglog,enumerator_id)

```

Note that `enumerator_id` is a data variable passed _without quotes_.

### Attaching cleaning log entries to datasets

In "reverse" of the above, we can add information from the cleaninglog to the dataset as "check" columns, indicating for each row wether a certain issue has been found, or whether a certain type of change was made. Each cleaning log check name now becomes a new logical column in the dataset that is set to `TRUE` for all rows where the cleaning log contained a matching row with an issue with that name:

```{r}
clog_cleaninglog_to_data(my_dataset,my_cleaninglog)

```

In this case there is only one 'type' of issue (called "manual inspection")
